More Minutes, More Points — The Role of Aggressive Play in NBA Scoring Trends

### 2024-04-19 paper/paper.qmd
last update

---
title: "More Minutes, More Points — The Role of Aggressive Play in NBA Scoring Trends"
subtitle: "An Analysis of NBA Player Performance Using Lasso Regression for the 2023-2024 Season"
author: 
  - Yihang  Cai
thanks: "Code and data are available at: https://github.com/peachvegetable/NBA-player-points"
date: today
date-format: long
abstract: "This paper investigates the predictive capacity of lasso regression in determining NBA players' total points from season statistics. The study found that minutes played, turnovers, and personal fouls are significant predictors of scoring, with turnovers unexpectedly indicating a higher involvement in gameplay leading to increased scoring. The implications of these findings extend beyond sports analytics, suggesting that player evaluations in team sports may benefit from a deeper understanding of player involvement as opposed to conventional performance metrics. By refining predictive models and reassessing performance indicators, this research offers new insights into effective player performance assessment and strategic planning in professional basketball."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

### woid space setup ###
library(tidyverse)
library(arrow)
library(tidymodels)
library(modelsummary)
library(ggplot2)
library(knitr)
library(janitor)
```

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-top5players
#| tbl-cap: Top 5 NBA Players Based on Select Predictors Highly Correlated with Points Scored, Season 2023-2024

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-top10players
#| tbl-cap: Top 10 NBA Players with selectely statistics, Season 2023-2024

@@

# show the first 10 players, and then rename some of the variables for a better view, and set the digits to two d.p. for a better view
analysis_data |>
  head(10) |>
  rename(
    'fg%' = fg_percent,
    'x3p%' = x3p_percent,
    'x2p%' = x2p_percent,
    'efg%' = e_fg_percent,
    'fg%' = fg_percent,
    'ft%' = ft_percent,
    'pos' = broad_position
  ) |>
  kable(digits = 0)
```

@@

weaknesses: position grouping; discussion: less important variables, like blk not served as scorer, age

>>>>

---
title: "More Minutes, More Points — The Role of Aggressive Play in NBA Scoring Trends"
subtitle: "An Analysis of NBA Player Performance, 2023-2024 Season"
author: 
  - Yihang  Cai
thanks: "Code and data are available at: https://github.com/peachvegetable/NBA-player-points"
date: today
date-format: long
abstract: "This paper investigates the predictive capacity of lasso regression in determining NBA players' total points from season statistics. The study found that minutes played, turnovers, and personal fouls are significant predictors of scoring, with turnovers unexpectedly indicating a higher involvement in gameplay leading to increased scoring. The implications of these findings extend beyond sports analytics, suggesting that player evaluations in team sports may benefit from a deeper understanding of player involvement as opposed to conventional performance metrics. By refining predictive models and reassessing performance indicators, this research offers new insights into effective player performance assessment and strategic planning in professional basketball."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

### woid space setup ###
library(tidyverse)
library(arrow)
library(tidymodels)
library(ggplot2)
library(knitr)
library(janitor)
```

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-top5players
#| tbl-cap: "Top 5 NBA Players by Predictors Highly Correlated with Points, 2023-2024 Season"

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-top10players
#| tbl-cap: Top 10 NBA Players by selected statistics, Season 2023-2024

@@

# show the first 10 players, and then rename some of the variables for a better view, and set the digits to two d.p. for a better view
analysis_data |>
  head(10) |>
  rename(
    'fg%' = fg_percent,
    'x3p%' = x3p_percent,
    'x2p%' = x2p_percent,
    'efg%' = e_fg_percent,
    'ft%' = ft_percent,
    'pos' = broad_position
  ) |>
  kable(digits = 1)
```

@@
